---
title: "Simple R notebook for OmniPathR in COVID-19"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r}
if(!'OmnipathR' %in% installed.packages()[,"Package"]){
    require(devtools)
    install_github('saezlab/OmnipathR')
}

require(OmnipathR)
```
```{r}
require(dplyr)
require(tibble)
require(purrr)
require(openxlsx)
require(igraph)
require(qgraph)
```
We download the OmniPath interactions
```{r}
ia_omnipath <- import_Omnipath_Interactions() %>% as_tibble()
```
```{r}
PRKACA_interactors<- ia_omnipath[ia_omnipath$source=="P17612",]
#write.table(head(PRKACA_interactors), "sample_interactions.csv", sep=",")
length(unique(PRKACA_interactors$target_genesymbol))
head(PRKACA_interactors)
```
We can also see various annotations of the proteins. E.g localisation or pathways.
Let's see what kind of databases are available.
```{r}
get_annotation_databases()
```
Let's check the cytosol proteins where the CAMP dependent protin kinase targets should be. We use the human protein atlas for that.
```{r}
PRKACA_targets_intracell_localisation <- import_Omnipath_annotations(select_genes= PRKACA_interactors$target_genesymbol, filter_databases=c("HPA_subcellular")) %>% as_tibble()
cytosol_portiens <- PRKACA_targets_intracell_localisation[PRKACA_targets_intracell_localisation$value=="Cytosol",]
```
Let'ss ay we want to sewee which proteins will be affected by PRKACA in the lung.
```{r}
PRKACA_targets_tissue_localisation <- import_Omnipath_annotations(select_genes= PRKACA_interactors$target_genesymbol, filter_databases=c("HPA_tissue")) %>% as_tibble()
```
```{r}
head(PRKACA_targets_tissue_localisation)
```
```{r}
organ_annotations <- PRKACA_targets_tissue_localisation[PRKACA_targets_tissue_localisation$label=="organ", ]
lung_PRKCA_protein_annotations <- organ_annotations[organ_annotations$value=="lung",]
records <- unique(lung_PRKCA_protein_annotations$record_id)

PRKACA_targets_tissue_localisation_lung_data <- PRKACA_targets_tissue_localisation %>% filter(record_id %in%c(records))
to_select_level <- PRKACA_targets_tissue_localisation_lung_data[PRKACA_targets_tissue_localisation_lung_data$label == "level", ]
record_ids_for_selection = to_select_level[to_select_level$value!="Not detected", "record_id"]
PRKACA_targets_expressed_in_lung <- PRKACA_targets_tissue_localisation_lung_data   %>% filter(record_id %in%c(record_ids_for_selection$record_id))
result<-PRKACA_targets_expressed_in_lung[PRKACA_targets_expressed_in_lung$label==c("tissue", "level"), ]
length(unique(PRKACA_targets_expressed_in_lung$uniprot))
head(result)
```
```{r}
write.table(lung_PRKCA_protein_interactors, "PRKCA_interactors_in_the_lung_HPA.csv", sep=",")
length(unique(lung_PRKCA_protein_interactors$uniprot))
```
Transcriptional regulation
If you want more interactions, pass the argument
confidence_level = c('A', 'B', 'C', 'D')
For descriptionof confidence levels see the paper: https://genome.cshlp.org/content/29/8/1363

```{r}
ia_transcriptional <- import_TFregulons_Interactions(confidence_level = c('A', 'B', 'C')) %>% as_tibble()
```
Let's say we want to find some transcription factors which regualte the ACE2 receptor
```{r}
ia_transcriptional[ia_transcriptional$target=="Q9BYF1",c("source", "source_genesymbol")]
```
